Investigating soundscapes perception through acoustic scenes simulation.

### 2017-12-13 lafaySimulatedScenePerception.tex
up until results

All the participants start by designing the ideal environment; they read the second set of instruction at the end of the first experiment. Participants are completely free of their choices concerning sounds and synthesis parameters. They must nevertheless respect the two following constraints :

@@

These constraints are notified in the instructions; no control is done \emph{a priori} in the simulation user interface.

@@

Each simulation process comprises two parts:

\begin{enumerate}
\item the simulation: this step is split in three steps:
\begin{itemize}
  \item  select the class of sounds
	\item  give a name to the selected class of sounds
	\item  set the parameters of the tracks related to the selected class of sounds, see Section~\ref{sec:simscene_parametre}.
\end{itemize}
\item the production of a free comment of the composed soundscape
\end{enumerate}

@@

In addition, once the two sound scenes are done, participants are invited to:

@@

Before starting the first simulation, a 20-minute tutorial is given to the participants in order to familiarize them with the simulation interface and the sound database. The experiment is planned to last about two hours and a half, including breaks that could be taken by the participants.

@@

All the participants do the experiment on desktop computers with the same hardware and software configurations. The audio files are played in diotic conditions by headphones. During the tutorial, participants are asked to adjust the global sound volume to a comfortable level. Once set, they are not allowed to modify it in the remaining of the experiment.

@@

All the participants do the experiment at the same time. They are equally split in three identical quiet rooms. They are not allowed to talk to each other during the experiment.

@@

Three experimenters (one in each room) are present during the whole duration of the experiment in order to assist participants with potential hardware and software issues and to answer questions.

@@

44 students (14 females; averaging $21.6$ years of age, s.d. of $2.0$ years) from \emph{Ecole Centrale de Nantes} (a french engineering school) took part to the experiment. All the subjects live in Nantes, France, for at least two years

@@

Among the 44 participants, 40 succeeded, producing at the end 80 simulated sound scenes (40 ideal scenes, 40 non ideal scenes). the 4 other participants were excluded from the process due to a lack of respect or understanding of the instructions, or for exceeding the maximum duration of the experiment.

@@

The assessment is done with a 7-point bipolar semantic scale going from -3 (non ideal / unpleasant) to +3 (ideal / pleasant). Before evaluating a scene, participants must listen to the first 20 seconds of the stimuli. After the evaluation, they are free to switch to the next scene.

@@

For each participant, sound scenes are played in a random order. The first 10 scenes allow the participants to calibrate their scores; as they are always composed of 5 ideal scenes and 5 non ideal scenes. These first 10 scenes are played back again at the end of the experiment and the last evaluations are the ones that are only taken into account in the results.

@@

All the participants do the experiment on computers with equal hardware and software configurations. The audio files are played in diotic conditions by semi-open headphones \emph{Beyerdynamic DT 990 Pro}. The stimuli are the scenes obtained by the simulation experiment of experiment 1.a. The output sound level is the same for all the participants.

@@

All the participants do the experiment simultaneously. They are not allowed to talk to each other during the experiment.

@@

An experimenter is present during the whole duration of the experiment in order to assist participants and possibly answer questions.

@@

A set of features is attached to each sound scene. The analysis is done on the basis of these features. A summary of these features (and the corresponding acronyms) is presented in Table ~\ref{tab:acronyme}. In order be consistent with the evaluation of experiment 1.b, features are not computed in the whole duration of the sequences but only on their 30-second reduced version used as stimuli for experiment 1.b, see Section~\ref{sec:xp1b_plan}.

@@

\item \emph{semantic features}: the sound classes that are present / absent from the scene. We use a boolean vector $S = (x_1, x_2, \ldots, x_n)$ that indicates the class of sounds involved in the scene. Each boolean $x$ of this vector corresponds to a specific class of sounds: $x = 1$ if the class is present in the scene, and $x = 0$ otherwise. The vector dimension ($n$) depends on the level of abstraction that is considered for analysis. For instance, for the abstraction level 1 that includes 44 classes of sounds, $n = 44$.

\item \emph{structural features}: the \emph{sound level}. To figure out the sound level, we draw inspiration from the $L_{Aeq}$ feature. In our case, it is a scalar computed from the  signal (in Volt and not in Pascal), and converted in decibels, with a reference of 1 Volt (full scale). The level is obtained by computing the quadratic mean of the signal every second and averaging the results over the total duration of the scene. An A-filtering module processes the data before the quadratic means are computed. We note $L$, $L(E)$ and $L(T)$ the computed levels by respectively considering the whole set of samples, only the set of event samples and only the set of texture samples.
\end{itemize}

@@

\item \emph{analysis of the perceived pleasantness}: the goal is to evaluate whether the perceived pleasantness differ between the i- and ni-scenes. We look at the existence of differences between the average $\mathcal{A}_{scene}$ and the average $\mathcal{A}_{subject}$ for each type of scenes (i or ni). The significance is evaluated by a two-sample Student test for $\mathcal{A}_{scene}$ and by a paired-sample Student test for $\mathcal{A}_{subject}$;

>>>>

All the participants start by designing the ideal environment; they read the second set of instructions at the end of the first experiment. Participants are completely free of their choices concerning sounds and synthesis parameters. The created sound environments must nevertheless fullfill the two following constraints :

@@

These constraints are notified in the instructions; no control is done \emph{a priori} in the simulation interface.

@@

Each simulation process is decomposed in several steps:

\begin{enumerate}
\item the simulation: the user is asked to:
\begin{itemize}
  \item  select the class of sounds
	\item  give a name to the selected class of sounds
	\item  set the parameters of the tracks related to the selected class of sounds, see Section~\ref{sec:simscene_parametre}.
\end{itemize}
\item the feedback : writing of a free form comment about the composed soundscape
\end{enumerate}

@@

In addition, once the two sound scenes are completed, participants are invited to:

@@

Before starting the first simulation, a 20-minute tutorial is given to the participants in order to familiarize them with the simulation interface and the sound database. The experiment is planned to last about two hours and a half, including breaks that the subjects are allowed to take.

@@

All the participants perform the experiment on standard desktop computers with the same hardware and software configurations. The audio files are played in diotic conditions using headphones. During the tutorial, participants are asked to adjust the sound volume to a comfortable level. Once set, they are not allowed to modify it during the remaining of the experiment.

@@

All the participants performed the experiment at the same time. They are equally distributed in three identical quiet rooms. They are not allowed to talk to each other during the experiment.

@@

Three experimenters (one in each room) are available during the whole duration of the experiment in order to assist participants with potential hardware and software issues and to answer questions.

@@

44 students (14 females; averaging $21.6$ years of age, s.d. of $2.0$ years) from \emph{Ecole Centrale de Nantes} (a french engineering school) took part to the experiment. All the subjects live in Nantes, France, for at least two years.

@@

Among the 44 participants, 40 succeeded, producing at the end 80 simulated sound scenes (40 ideal scenes, 40 non ideal scenes). the 4 other participants were excluded from the process due to a lack of understanding of the instructions or failure to respect them, or for exceeding the maximum duration allowed to perform the experiment.

@@

The assessment is done with a 7-point bipolar semantic scale going from -3 (non ideal / unpleasant) to +3 (ideal / pleasant). Before evaluating a scene, participants must listen to the first 20 seconds of the stimuli. After the evaluation, they are free to continue to the next scene.

@@

For each participant, sound scenes are played in a quasi random order. 5 ideal scenes and 5 non ideal scenes are first sequenced to allow the participants to calibrate their scores. These first 10 scenes are played back again at the end of the experiment. Only the last evaluations are taken into account.

@@

All the participants perform the experiment on computers with equivalent hardware and software configurations. The audio files are played in diotic conditions by semi-open headphones \emph{Beyerdynamic DT 990 Pro}. The stimuli are the scenes obtained by the simulation experiment of experiment 1.a. The output sound level is the same for all the participants.

@@

All the participants perfrom the experiment simultaneously. They are not allowed to talk to each other during the experiment.

@@

An experimenter is available during the whole duration of the experiment in order to assist participants and answer questions any questions the participants may have.

@@

A set of features is attached to each sound scene. The analysis is done on the basis of these features. A summary of these features (and the corresponding acronyms) is presented in Table ~\ref{tab:acronyme}. In order to be consistent with the evaluation of experiment 1.b, features are not computed in the whole duration of the sequences but only on their 30-second reduced version used as stimuli for experiment 1.b, see Section~\ref{sec:xp1b_plan}.

@@

\item \emph{semantic features}: We use a boolean vector $S = (x_1, x_2, \ldots, x_n)$ that indicates the class of sounds involved in the scene, \textit{i.e.} the sound classes that are present / absent from the scene. Each boolean $x$ of this vector corresponds to a specific class of sounds: $x = 1$ if the class is present in the scene, and $x = 0$ otherwise. The vector dimension ($n$) depends on the level of abstraction that is considered for the analysis. For instance, for the abstraction level 1 that includes 44 classes of sounds, $n = 44$.

\item \emph{structural features}: the \emph{sound level}. To figure out the sound level, we draw inspiration from the $L_{Aeq}$ feature. In our case, it is a scalar computed from the  signal (in Volt and not in Pascal), and converted in decibels, with a reference of 1 Volt (full scale). The level is obtained by computing the quadratic mean of the signal every second and averaging the results over the total duration of the scene. An A-filtering module processes the data before the quadratic means are computed. We note $L$, $L(E)$ and $L(T)$ the computed levels by respectively considering the whole set of samples, only the set of event samples, and only the set of texture samples.
\end{itemize}

@@

\item \emph{analysis of the perceived pleasantness}: the goal is to evaluate whether the perceived pleasantness is in accordance with the pleasantness label given by the creators of the  i- and ni-scenes durint the experiment 1.a. To do so, we consider the differences between the average $\mathcal{A}_{scene}$ and the average $\mathcal{A}_{subject}$ for each type of scenes (i or ni). The significance is evaluated by a two-sample Student test for $\mathcal{A}_{scene}$ and by a paired-sample Student test for $\mathcal{A}_{subject}$;
